### 2018-08-12 extended_abstract/improvements.tex
latest

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%     File: Thesis_Model_Improvements.tex                              %
%     Tex Master: Thesis.tex                                           %
%                                                                      %
%     Author: João C. Godinho                                          %
%     Last modified : Apr 2018                                         %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Model Improvements}
\label{chapter:model_improvements}

Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this chapter to present the improvements made in order to obtain a more robust model to detect malware.
We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Multi Layer Model}
\label{section:improvements_multi_layer}

On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.
Although that was the end purpose of our work, it does not give any deeper understanding of the sample, other than being or not malware.
With that in mind, we propose a new model that not only provides the probability of a sample being malware, but also the probability of being from a specific malware type.

This new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_multilayer.png}
	\caption{Multi layer model representation.}
	\label{fig:dia_multilayer}
\end{figure}

@@

With this new model defined, we now present our approach on selecting the $n$ classes of interest.
This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.

To choose the malware classes for our datasets, we take into account our description in Chapter \ref{chapter:related_work} regarding propagation methods and malware names based on purpose.
With this in mind, we chose 6 malware classes: \textit{virus}, \textit{trojan}, \textit{worm}, \textit{ransom}, \textit{spyware} and \textit{other}.
The first three classes, \textit{virus}, \textit{trojan} and \textit{worm}, were chosen from the propagation methods, whereas \textit{ransom} and \textit{spyware}
where chosen due to their popularity in recent years. The last class, \textit{other}, serves for any malware that does not fit the previous five.

As mentioned in Chapter \ref{chapter:related_work}, there is no agreed upon naming convention for malware, which translates into different names for the same malware sample.
To minimize this problem, we referenced a tool by Sebastián, M. et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, using the names provided by VirusTotal~\cite{tool:virustotal}.

We took advantage of this tool and modified it such that instead of providing a family name, it would provide one (or more) of the 6 previously defined classes.
Specifically, we changed it in a way that given a set of malware names, the output would be a distribution over the 6 malware classes.

@@

where $f_c$ is the frequency for the class $c$ and $\sum_{c}f_c$ is the number of times all classes appear.
For example, if a given set of names contain the name \textit{trojan} 3 times and the name \textit{virus} one time, then the weights would be
\begin{eqnarray*}
	\mathcal{W}_{trojan}=\dfrac{3}{4}=0.75,~\mathcal{W}_{virus}=\dfrac{1}{4}=0.25,~ \mathcal{W}_{c}=0, c \in \{worm, spyware, other, ransom\}
\end{eqnarray*}

@@

Having these malware classes defined for our multi layer model, we also added the \textit{goodware} class for samples that are not malware.
Doing so gives us 7 possible classes, 6 of which are malware only.
It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise, if it belongs to any malware class, it cannot belong to the \textit{goodware} class.

The following Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.
Note that the classes are not exclusive, as such the represented numbers show how many samples, out of each dataset $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$, are classified by at least one vendor as belonging to any of the defined classes.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Trojan			& 97,054 & 44,329 & 24,176\\
		Other			& 49,443 & 24,126 & 14,750\\
		Worm			& 24,554 & 14,837 & 9,381\\
		Virus			& 21,055 & 12,531 & 6,899\\
		Spyware			& 20,724 & 10,172 & 5,955\\
		Ransom			& 7,761 & 1,924 & 1,160\\
		\bottomrule
	\end{tabular}
	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

One can see that the \textit{trojan} class is present in almost all samples, whereas \textit{ransom} is the least common class.
More detailed statistics regarding each dataset is presented in Appendix \ref{chapter:appendixClasses}, containing information like the average weight for each class and the minimum and maximum values.

@@

In sum, we added a layer of labeling to our dataset, where a sample can either belong to the \textit{goodware} class, or to a set of the other 6 malware classes.
This concludes the description of our first improvement to the model, what follows are our improvements regarding features.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Dynamic Features}
\label{section:improvements_dynamic_features}

Having described our first model improvement, we now proceed to our second enhancement, which takes advantage of the amount of information available in Cuckoo's~\cite{tool:cuckoo} reports.

In the previous chapter, we used static imports as features for our malware detection model.
Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
To overcome these limitations, we resort to more dynamic information provided by Cuckoo.

Given that Cuckoo runs the provided samples inside a virtual machine, it monitors the sample from a dynamic point of view.
This monitoring includes information like the sequence of library calls and their categories and network and file activity.
With this in mind, we now detail what dynamic information we extracted from the reports.

@@

Figure \ref{fig:boxplot_category_calls} shows a boxplot for the category calls, where one can see the categories \textit{filesystem}, \textit{registry} and \textit{system} are the ones with higher mean, hence more used than the other categories.
Although not depicted, some outliers for the number of category calls go up to 1,400,000, which do not provide useful information other than the number of category calls is very high.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=\textwidth]{Figures/boxplot_category_calls.png}
	\caption{Category calls boxplot.}
	\label{fig:boxplot_category_calls}
\end{figure}

With that in mind, we decide to transform the values to follow a normal distribution.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.

@@

Our second type of dynamic information are the number of library calls, which closely relate to the previous feature.
While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cuckoo Signatures}
\label{section:improvements_categories}

@@

Although some signatures are defined by a sequence of library calls, which could also be attained by creating \textit{n-grams} of our previously defined feature, using the already defined Cuckoo signatures saves training time.
As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, which becomes highly unpractical when considering a higher degree of n-grams.

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Malware			& 94,248 & 42,911 & 23,437\\
		Goodware		& 28,385 & 1,741 & 1,741\\
		\midrule
		Total			& 122,633 & 44,652 & 25,178\\
		\bottomrule
	\end{tabular}
	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_improved}
\end{table}

@@

In the previous sections we described all our modifications to the original model $\LR$: a two layer approach that splits malware into subclasses and the addition of three dynamic features.
We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.

Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.

We again use the resources provided by INOV - Inesc Inovação, together with scikit-learn~\cite{tool:sklearn}, Pandas~\cite{tool:pandas} and Jupyter Interactive Notebooks~\cite{tool:jupyter} to train and analyze the data.

Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
For this scenario, we obtain an \gls{auroc} of 98\%, as presented by the red curve in Figure \ref{fig:improved_xval}.
These provide the highest results, provided by factors like a small and reliable dataset, more features and the use of cross-validation.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/xval_improved.png}
	\caption{Multi layer results for dynamic features in laboratory conditions.}
	\label{fig:improved_xval}
\end{figure}

@@

As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/pastpresent_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
		fold 1)}
	\label{fig:pastpresent_modified}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/presentpast_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
	\label{fig:presentpast_improved}
\end{figure}

@@

Finally, we retest how a reduced training set behaves in our scenarios by using our \textit{Temporal Window} methodology.
As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $w=3$ as in Section \ref{section:single_layer_results}.
In summary, we selected 30\% of each dataset for training and 10\% for validation.
Starting at the oldest fold, we apply this window and slide it by one fold at each iteration.
Figure \ref{fig:slidingwindow_modified} shows how all our scenarios $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$ score the same \gls{auroc} of 94\%, although with different curves.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:slidingwindow_modified}
\end{figure}

@@

As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
The improved layer results are summarized in Table \ref{tab:multilayer_results}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{ccccc}
		\toprule
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\midrule
		Cross-Validation & 0.98 & 0.98 & 0.95 & 90 / 10\\
		Past-to-Present & 0.96 & 0.96 & 0.92 & 10 to 90 / 10\\
		Present-to-Past & 0.97 & 0.98 & 0.96 & 10 to 90 / 10\\
		Sliding-Window & 0.94 & 0.94 & 0.94 & 30 / 10\\
		\bottomrule
	\end{tabular}
	\caption{Multi layer results summary.}
	\label{tab:multilayer_results}
\end{table}

Having studied how our modifications improved the proposed model, the following chapter goes on to provide practical applications for our work.

>>>>

%!TEX root = main.tex

\section{Model Improvements}
\label{sec:improvements}

Having a solid baseline model for our malware detection task together with how laboratory \textit{vs.}\ real-world scenarios change the model outcome, we now take this section to present the improvements made in order to obtain a more robust model to detect malware.
We start by describing our first improvement, applying a multi layer model to extract more information regarding a sample.
We then take this enhanced model and increase the number of features to include dynamic content and how it impacted the model's results.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multi Layer Model}
\label{section:improvements_multi_layer}

On the previous chapter we ended up with a simple \gls{lr} model $\LR$ that given a set of static imports from a sample, would give the probability of it being malware.

In this section we provide a new model $\mathcal{E}$ comprises a simple ensemble stacking approach, which instead of using a single \gls{lr} classifier, multiple ones are used, layered into two steps.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=\columnwidth]{dia_multilayer}
	\caption{Multi layer model representation.}
	\label{fig:dia_multilayer}
\end{figure}

@@

We now present our approach on selecting the $n$ classes of interest.
This represents another labeling problem, but now instead of having to label between goodware and malware, we have to label the malware as belonging to some subclass.

With this in mind, we chose 6 malware classes: \textit{virus}, \textit{trojan}, \textit{worm}, \textit{ransom}, \textit{spyware} and \textit{other}.

To help label our malware samples into the aforementioned classes, we take a tool by Sebastián, M. et al.~\cite{sebastian2016avclass}, AVClass, which was built to normalize a malware sample name into the most likely family, and modify it such that instead of providing a family name, it would provide one (or more) of the 6 previously defined classes.
Specifically, we changed it in a way that given a set of malware names, the output would be a distribution over the 6 malware classes.

@@

where $f_c$ is the frequency for the class $c$ and $\sum_{c}f_c$ is the number of times all classes appear.
For example, if a given set of names contain the name \textit{trojan} 3 times and the name \textit{virus} one time, then the weights would be

\begin{eqnarray*}
	\mathcal{W}_{trojan}=\dfrac{3}{4}=0.75,~\mathcal{W}_{virus}=\dfrac{1}{4}=0.25,\\\mathcal{W}_{c}=0, c \in \{worm, spyware, other, ransom\}
\end{eqnarray*}

@@

Having these malware classes defined for our multi layer model, we also added the \textit{goodware} class for samples that are not malware.
Doing so gives us 7 possible classes, 6 of which are malware only.
It is worth mentioning that if a sample belongs to the \textit{goodware} class, it cannot belong to any other, likewise, if it belongs to any malware class, it cannot belong to the \textit{goodware} class.
Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\hline
		Trojan			& 97,054 & 44,329 & 24,176\\
		Other			& 49,443 & 24,126 & 14,750\\
		Worm			& 24,554 & 14,837 & 9,381\\
		Virus			& 21,055 & 12,531 & 6,899\\
		Spyware			& 20,724 & 10,172 & 5,955\\
		Ransom			& 7,761 & 1,924 & 1,160\\
		\hline
		\hline
		Malware Total	& 98,582 & 45,306 & 24,658\\
		\hline
	\end{tabular}
	\medskip
	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

@@

In the base model, we used static imports as features for our malware detection model.
Although the results are reasonable, the information which can be retrieved from static imports alone is limited.
As an example, if a sample is compressed, encrypted or packed, its behavior cannot be inferred from static imports only.
To overcome these limitations, we resort to more dynamic information provided by Cuckoo.

@@

To normalize the category calls value, we decide to transform the values to follow a normal distribution.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.

@@

Our second type of dynamic information are the number of library calls.
While \textit{category calls} provide the number of calls for a given category, \textit{library calls} provide the count for each library call, hence being a subset of the previous.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cuckoo Signatures}
\label{section:improvements_signatures}

@@

@@

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\hline
		Malware			& 94,248 & 42,911 & 23,437\\
		Goodware		& 28,385 & 1,741 & 1,741\\
		\hline
		Total			& 122,633 & 44,652 & 25,178\\
		\hline
	\end{tabular}
	\medskip
	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_improved}
\end{table}

@@

We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{sec:eval_results}.

Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
\medskip

Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
For this scenario, we obtain an \gls{auroc} of 98\%, as presented by the red curve in Figure \ref{fig:improved_xval}.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{xval_improved}
	\caption{Multi layer results for dynamic features in laboratory conditions.}
	\label{fig:improved_xval}
\end{figure}

@@

As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are not affected at all, when the change between the scenarios is merely in the number of malware samples.
Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=\columnwidth]{pastpresent_improved}
	\caption[Multi layer results for dynamic features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
		fold 1)}
	\label{fig:pastpresent_modified}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{presentpast_improved}
	\caption[Multi layer results for dynamic features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
	\label{fig:presentpast_improved}
\end{figure}

@@

Finally, we retest how a reduced training set behaves in our scenarios by using our \textit{Temporal Window} methodology.
As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $n=3$ as in Section \ref{section:single_layer_results}.
Starting at the oldest fold, we apply this window and slide it by one fold at each iteration.
Figure \ref{fig:slidingwindow_modified} shows how all our scenarios $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$ score the same \gls{auroc} of 94\%, although with different curves.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=\columnwidth]{Figures/slidingwindow_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:slidingwindow_modified}
\end{figure}

@@

%As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
The improved layer results are summarized in Table \ref{tab:multilayer_results}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{ccccc}
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\hline
		Cross-Validation & 0.98 & 0.98 & 0.95 & 90 / 10\\
		Past-to-Present & 0.96 & 0.96 & 0.92 & 10 to 90 / 10\\
		Present-to-Past & 0.97 & 0.98 & 0.96 & 10 to 90 / 10\\
		Sliding-Window & 0.94 & 0.94 & 0.94 & 30 / 10\\
		\hline
	\end{tabular}
	\medskip
	\caption{Multi layer results summary.}
	\label{tab:multilayer_results}
\end{table}

### 2018-08-12 thesis_report/Thesis_Background.tex
latest

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cuckoo Sandbox and Malwr}
\label{section:cuckoo}

@@

Some of the defined sections are not always available.
Their availability depends both on a successful analysis by Cuckoo, as well as the type of submitted file.
For example, only binary executables (\eg\ \gls{dll}, \gls{pe}) provide static analysis information.

@@

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO mais umas linhas sobre estatico e dinamico
\section{Cuckoo Sandbox and Malwr}
\label{section:cuckoo}

@@

Some of the defined sections are not always available.
Their availability depends both on a successful analysis by Cuckoo, as well as the type of submitted file.
For example, only binary executables (\eg\ \gls{dll}, \gls{pe}) provide static analysis information.
Using the defined sections provided by Cuckoo's reports, one can obtain static and dynamic features from the samples.

@@

\subsection{Best Practices for \gls{ml} in Malware Detection}

@@

\subsection{Malware Detection using Static and Dynamic Features}

@@

\subsection{Pitfalls of Using Classical Validation Methodologies}

### 2018-08-12 thesis_report/Thesis_Conclusions.tex
latest

@@

\begin{itemize}
	\item We provided means to achieve metrics as to facilitate the malware labeling task.
	\item We developed methodologies that test real-world scenarios which take into account time consistency.
	\item We provided a simple model to detect malware and how it behaves under different methodologies.
	\item We improved the previous model to boost the results in all of our tested methodologies.
	\item We started developing practical applications of our model, to be inserted in real world scenarios.
\end{itemize}

@@

As future work, the developed model could be improved by adding more features, and by doing better research regarding the impacts of each feature.
Other models could be tested, as well as used in conjunction with our logistic regression, to improve the overall results.
Diverging from a supervised learning methodology to include other types of learning would also bring benefits and possibly achieve a fully automated pipeline to detect malware, with minimal user interaction.

>>>>

% TODO mudar pratical para aqui

\section{Practical Applications}
\label{section:practical_applications}

We take this section to briefly describe how we took the previously defined model to the real world, providing practical implementations for it.
The following work is still on-going as part of a scholarship provided by INOV - Inesc Inovação, as such the end products are still being developed and test.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Malware Detection Service}
\label{section:malware_service}

Our first practical application for the obtained model comes as a malware detection service.
The idea is to take advantage of Cuckoo~\cite{tool:cuckoo} to implement a web service identical to Malwr~\cite{tool:malwr}, but also adding the malware detection by applying our model.

To implement this service, we developed a simple web application that receives files from users and then sends them to a Cuckoo instance running on a server.
We created a custom signature for Cuckoo that takes the analysis results and converts it into features suitable for our model, which is then used to classify the sample.
The output of the model is then appended as a signature to the final report.
After analysis our web application extracts basic information regarding the file, as well as the output of the classifier.

We implemented our web application using Django~\cite{tool:django}.
The implemented application provides a simple upload form for users to submit their own files.
Upon submission, Django sends the sample to a Cuckoo instance running locally.
While Cuckoo is analyzing the sample, users are redirected to the report page, which shows the current state of analysis.
Upon finishing analysis, the page changes to yield the classifier results.

Figure \ref{fig:malware_detetor} shows the service being tested with a \texttt{petya}\footnote{Symantec - Petya Ransomware, October 27, 2017 [https://www.symantec.com/blogs/threat-intelligence/petya-ransomware-wiper]} sample, using early versions of our $\LR$ model.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/malware_detector.png}
	\caption{Malware Detector in early stages.}
	\label{fig:malware_detetor}
\end{figure}

We are currently testing and optimizing the service, as it is not currently using our latest and best model $\mathcal{E}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Email Attachment Scanner}
\label{section:email_scanner}

Our second practical application uses the previous service, but in an automated way, to scan email attachments arriving and departing INOV's mail server.
The idea is to insert the malware detector in the mail server pipeline as a scanner.

INOV's mail server uses MailScanner~\cite{tool:mailscanner} to filter and scan emails.
In Figure \ref{fig:mailscanner} we provide a diagram on the MailScanner pipeline.

Our implementation works by creating a custom \textit{Virus Scanner}, which takes email attachments and sends them to the previous described malware detection service.
Upon receiving the results, our custom scanner modifies the mail object to include the classification results.

Further down in the \textit{Message Processing} step, we added a custom step to modify an unused \texttt{Header} in the email to include the classification results.

Currently we are in the process of moving the developed code into the mail server to start obtaining real world results.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/mailscanner.png}
	\caption{MailScanner process diagram.}
	\label{fig:mailscanner}
\end{figure}

@@

\begin{itemize}
	\item We provided means to achieve metrics as to facilitate the malware labeling task.
	\item We developed methodologies that test real-world scenarios which take into account time consistency, rather than laboratory created scenarios.
	\item We provided a simple model to detect malware and investigated how it behaves under different methodologies.
	\item We improved the previous model with a two layer approach and more dynamic features to boost the results in all of our tested methodologies.
	\item We started developing practical applications of our model, to be inserted in real world scenarios.
\end{itemize}

@@

As future work, the developed model could be improved by adding more features, and by doing more in depth research regarding the impacts of each feature.
Other models could be tested, as well as used in conjunction with our logistic regression, to improve the overall results.
Diverging from a supervised learning methodology to include other types of learning would also bring benefits and possibly achieve a fully automated pipeline to detect malware, with minimal user interaction.

### 2018-08-12 thesis_report/Thesis_Introduction.tex
latest

The amount of malware has shown an exponential increase since the early 2000s, from 1 million in 2006 to over 500 million by the end of 2016~\cite{av-test:report,mcafee:report}. According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http:\/\/www.newsfactor.com\/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites. A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results. From the contact with malicious software, 1.1 billion identities (\ie\ user's personal information) have been exposed in 2016 alone~\cite{symantec:report}.

@@

@@

We also provide several improvements to the model.
One of these come under the form of a multi-layer approach, which instead of classifying a sample as malware or goodware, also provides information regarding the type of malware.
Our other improvement regards providing more information about a sample to the model, in order to obtain higher scores.

@@

Chapter \ref{chapter:static_features} and \ref{chapter:model_improvements} takes the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.

Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service and how it was inserted into a corporate email server to analyze attachments.

Chapter \ref{chapter:conclusions} provides some final remarks on our work and contributions, together with possible ideas that may improve the presented work.

>>>>

The amount of malware has shown an exponential increase since the early 2000s, from 1 million in 2006 to over 500 million by the end of 2016~\cite{av-test:report,mcafee:report}. According to a 2007 study by McAfee\footnote{McAfee Study Finds 4 Percent of Search Results Malicious, by Frederick Lane, June 4, 2007 [http://www.newsfactor.com/story.xhtml?story\_id=010000CEUEQO]} about 4\% of the query results by major search engines lead to potentially dangerous websites. A more recent study by Norton in 2010\footnote{Symantec Press Release, July 29, 2010 [https://www.symantec.com/en/aa/about/newsroom/press-releases/2010/symantec\_0729\_02]} showed that at least 10\% of top trending search terms returned malicious results. From the contact with malicious software, 1.1 billion identities (\ie\ user's personal information) have been exposed in 2016 alone~\cite{symantec:report}.

@@

By designing these three scenarios, we can better understand how having a smaller and very distinctive dataset compares to a bigger and more uneven dataset, providing a greater understanding of the effects of \textit{ground-truth}.

@@

This will help us understand if one should take precautions when using classical evaluation methodologies, as these do not take into account time dependency between samples.

We also provide several improvements to the model.
One of these come under the form of a multi-layer approach, which instead of classifying a sample as malware or goodware, also provides information regarding the type of malware.
Our other improvement regards providing more information about a sample to the model, by using dynamic information extract during runtime, in order to obtain higher scores.

@@

Chapter \ref{chapter:static_features} and \ref{chapter:model_improvements} take the created datasets to build a malware detection model based on both static and dynamic features, as well as how these perform under different validation methodologies.

% Chapter \ref{chapter:practical_applications} describes how our model is practically implemented both as an online malware detection service and how it was inserted into a email server to analyze attachments.

Chapter \ref{chapter:conclusions} provides some final remarks on our work, how our model is practically implemented, contributions and possible ideas that may improve the presented work.
